---
title: "Covid Prediction"
author: "Alex Olmeta, Alex Paine, Donald Stricklin"
format: revealjs
embed-resources: true
editor: visual
---

## Project Overview

```{r}
#| echo: false
#| output: false
library(tidyverse)
library(readr)
library(kableExtra)

covid <- read_csv("data/owid-covid-data.csv") %>% 
  janitor::clean_names()


unique(covid$location)

```

::: {.fragment .fade-in-then-semi-out}
-   The world was not ready for the severe impact of COVID-19.
:::

::: {.fragment .fade-in-then-semi-out}
-   7,010,568 Reported COVID-19 deaths (as of 12/31/23)
:::

::: {.fragment .fade-in-then-semi-out}
-   Models on disease growth at the time were poorly equipped to stifle the spread of this deadly disease
:::

::: {.fragment .fade-in-then-semi-out}
-   This project has the potential to evaluate where a country stands in terms of projected death rate, thereby reinforcing or introducing new methods to prevent the spread, and thus, preventing the loss of innocent lives.
:::

## Objectives

::: {.fragment .fade-in}
***To Create a Predictive Model to Predict the Number of Deaths of an Unseen Country***
:::

::: {.fragment .fade-in-then-semi-out}
-   Assess the model's predictive performance by evaluating its accuracy, precision, recall, and other relevant metrics.
:::

::: {.fragment .fade-in-then-semi-out}
-   Model has application for both future COVID-19 variants + future similar diseases
:::

::: {.fragment .fade-in-then-semi-out}
-   Our proposed model would take several weeks of prior data on a previously unseen country and predict the number of deaths in the coming weeks.
:::

## The Data

::: {.fragment .fade-in-then-semi-out}
-   Comprehensive scope of COVID-19 metrics, including cases, deaths, vaccinations, and ICU admissions for 200+ countries/territories
:::

::: {.fragment .fade-in-then-semi-out}
-   After initial preprocessing, have 255,000 observations and 67 variables
:::

::: {.fragment .fade-in-then-semi-out}
-   Significant missingness issues - total cases reported every week, not day.
:::

::: {.fragment .fade-in-then-semi-out}
-   Some countries also do not provide ICU and vaccination rates making them harder to work with
:::

## Visualizing the Amount of New Deaths

```{r}
#| output-location: slide 
df <- covid %>% 
  filter(
    !is.na(new_deaths),
    location %in% c("United States", "Canada", "France", "Germany", "United Kingdom")
  ) %>% 
  mutate(
    location = factor(location,
                      levels = c("United States", "United Kingdom", "Germany", "France", "Canada"))
  )
ggplot(df, aes(date, new_deaths, color = location)) +
  #geom_point(alpha = 0.8) + 
  geom_line(alpha = 0.8) + 
  theme_minimal() +
  labs(
    x = "",
    y = "Deaths",
    title = "New Covid Deaths since 2020",
    color = "",
    caption = "Source:  Our World in Data"
  )
```

## Cleaning and Preprocessing

::: {.fragment .fade-in-then-semi-out}
-   Data was already largely preprocessed and came in one neat `.csv` file
:::

::: {.fragment .fade-in-then-semi-out}
-   Had to remove \~20,000 rows of regional summary information (e.g., case data for Europe) as these were redundant
:::

::: {.fragment .fade-in-then-semi-out}
-   Fair amount of missing data will need to be imputed (daily vs. weekly)
:::

::: {.fragment .fade-in-then-semi-out}
-   Some predictors are correlated
:::

## Conclusion

::: {.fragment .fade-in-then-semi-out}
-   Develop model based on daily COVID-19 data from the last 4 years that predicts the number of deaths in a country
:::

::: {.fragment .fade-in-then-semi-out}
-   We anticipate fewer deaths all over the world in the near future as COVID-19 has been trending downward
:::

::: {.fragment .fade-in}
-   Challenges
    -   Predictor selection
    -   Missing data
    -   Computation time
:::

## Citations

-   World Health Organization 2023 data.who.int, WHO Coronavirus (COVID-19) dashboard \> Deaths \[Dashboard\]. <https://data.who.int/dashboards/covid19/deaths>

-   Mathieu, E., Ritchie, H., Ortiz-Ospina, E. et al. A global database of COVID-19 vaccinations. Nat Hum Behav (2021). <https://doi.org/10.1038/s41562-021-01122-8>
